title: "Project Notebook"
author: "Claire Zhao"
output:
  prettydoc::html_pretty:
    highlight: github
    toc: yes
    toc_depth: 2
    theme: architect
  '# prettydoc::html_pretty': default
  

# set-up
```{r}
rm(list = ls())     # clear the workspace 
library(ISLR)       # load ISLR data package
library(tidyverse)
library(ggplot2)
```


# import data（我在本地run的，在github得改成"../data/WA_Fn-UseC_-HR-Employee-Attrition.csv"）
```{r}
raw_data <-
  read.csv(
    "WA_Fn-UseC_-HR-Employee-Attrition.csv",
    stringsAsFactors = T,
    na.strings = '?'
  )
names(raw_data)[names(raw_data) == 'ï..Age'] <- 'Age'
glimpse(raw_data)
```
# pre

# Classification Trees

```{r}
library(rpart)
library(rpart.plot)
```

ct_model<-rpart(..................# model formula
                data=Default,                             # dataset
                method="class",                           # "class" indicates a classification tree model 
                control=rpart.control(cp=0.03,maxdepth=4))   # tree control parameters. 
```
