---
title: "Credit Card Default"
Author: "Claire Zhao"
output:
  word_document:
    toc: yes
  html_notebook:
    number_sections: yes
    toc: yes
---
#set-up
```{r}
library(rmarkdown)
if (!require(kableExtra)) install.packages("kableExtra")
library(kableExtra)
library(knitr)
library(magrittr)

library(tidyverse)
library(ggplot2)
library(rpart)
library(rpart.plot)

if (!require(e1071)) install.packages("e1071")
library(e1071)
if (!require(fastDummies)) install.packages("fastDummies")
library(fastDummies)

if (!require(tree))
  install.packages('tree')
library(tree)

```


# import data
```{r}
set.seed(1)
raw_data <-
  read.csv(
    "WA_Fn-UseC_-HR-Employee-Attrition.csv",
    stringsAsFactors = T,
    na.strings = '?'
  )
names(raw_data)[names(raw_data) == 'ï..Age'] <- 'Age'
glimpse(raw_data)
```

# prepare& build the tree
```{r}
tmp_Attrition <- raw_data$Attrition
data_droped_col <- subset(raw_data, select = -c(Attrition, EmployeeCount, StandardHours, Over18))
data_dummied <- dummy_cols(data_droped_col)
data_processed <-
  data_dummied[, sapply(data_dummied, class) != "factor"]
colnames(data_processed) <- gsub(" ", "", colnames(data_processed))
colnames(data_processed) <- gsub("-", "", colnames(data_processed))
colnames(data_processed) <- gsub("&", "", colnames(data_processed))
colnames(data_processed) <- gsub("`", "", colnames(data_processed))
data_processed$Attrition <- tmp_Attrition
glimpse(data_processed)

# split sample and test
train <- sample(1:nrow(data_processed), nrow(data_processed) / 2)
test <- data_processed[-train,]

#Decision Tree
# plant a tree
data.tree <- tree(Attrition ~ ., data = data_processed)

summary(data.tree)
#+ fig.width=11, fig.height=8
plot(data.tree)
text(data.tree, pretty = 0)


data.tree <- tree(Attrition ~ ., data =data_processed, subset = train)
data.tree.pred = predict(data.tree, test, type = "class")

table(data.tree.pred, test$Attrition)
mean(data.tree.pred == test$Attrition)
```


# cross validating different trees 
```{r}
cv.data <- cv.tree(data.tree, FUN = prune.misclass)
names(cv.data)
#+ fig.width=11, fig.height=8
plot(cv.data)
cv.data
```

# get the best tree
```{r}
prune.data <-prune.misclass(data.tree, best = cv.data$size[which.min(cv.data$dev)])
#+ fig.width=11, fig.height=8
plot(prune.data)
text(prune.data, pretty = 0)

prune.data.pred <- predict(prune.data, test, type = "class")
table(prune.data.pred, test$Attrition)
mean(prune.data.pred == test$Attrition)
```
#****************************************************************************************************************
#从这里开始是prof 的方法 ct_model和k-fold
```{r}
ct_model<-rpart(Attrition ~.,           # model formula 
                data=data_processed,                      # dataset
                method="class",                           # "class" indicates a classification tree model 
                control=rpart.control(cp=0))  
rpart.plot(ct_model)   # tree plot
```

#Get the predicted value - class membership (yes or no) --> using a cut-off of 50%. 
```{r}
ct_pred_class<-predict(ct_model,type="class") # class membership (yes or no) 
head(ct_pred_class)

ct_pred<-predict(ct_model)  # get the predicted values - class probabilities (default)
head(ct_pred)
```
```{r}
data_processed$ct_pred_prob<-ct_pred[,2]   
print(ct_model)        # model results 
```
```{r}
summary(ct_model)      # model result details 
```

#train/test split ：我们的sample size是不是只有1400？所以我觉得这里可能有点问题 but i can‘t figure out
```{r}
set.seed(1)   # set a random seed 
index <- sample(10000, 2000) # random selection of indices. 20%
index
```


#test & training dataset：这里的varibale being trained我按照之前model写的 不确定！！！！需要铭豪大佬！
```{r}
test <- data_processed[index,c("Attrition","Age","OverTime_No","MonthlyIncome","JobRole_SalesExecutive","MaritalStatus_Single","YearsSinceLastPromotion","TotalWorkingYears","JobRole_ResearchScientist","StockOptionLevel","JobSatisfaction","RelationshipSatisfaction")]       
# save 20% as a test dataset:index

training <-data_processed[-index,c("Attrition","Age","OverTime_No","MonthlyIncome","JobRole_SalesExecutive","MaritalStatus_Single","YearsSinceLastPromotion","TotalWorkingYears","JobRole_ResearchScientist","StockOptionLevel","JobSatisfaction","RelationshipSatisfaction")]   
# save the rest as a training set:-index 

test
training


```

### build a model with a training set. cp用了0.03
```{r}
training_model<-rpart(Attrition ~.,
                      data=data_processed, 
                      method="class", 
                      control=rpart.control(cp=0.03))

rpart.plot(training_model)
```
### predicting probabilities/class labels for test data 我这里开始做不下去了所以换了k-fold写

### k-fold Cross-validation，cp用了0

```{r}
set.seed(1)   # set a random seed 
full_tree<-rpart(Attrition~.,
                     data=training, 
                     method="class",
                     control=rpart.control(cp=0))  #cp=0, more complex, let the tree grow
rpart.plot(full_tree)
```
```{r}
printcp(full_tree)   # xerror, xstd - cross validation results  
```
```{r}
plotcp(full_tree)    
```

```{r}
min_xerror<-full_tree$cptable[which.min(full_tree$cptable[,"xerror"]),]
min_xerror

# prune tree with minimum cp value
min_xerror_tree<-prune(full_tree, cp=min_xerror[1])
rpart.plot(min_xerror_tree)
```
```{r}
bp_tree<-min_xerror_tree
test$ct_bp_pred_prob<-predict(bp_tree,test)[,2]
test$ct_bp_pred_class=ifelse(test$ct_bp_pred_prob>0.5,"Yes","No")

table(test$ct_bp_pred_class==test$Attrition)  # error rate

table(test$ct_bp_pred_class,test$Attrition, dnn=c("predicted","actual"))  # confusion table on test data
```
